Paper 1

### 2022-06-15 scripts/Paper 1.Rmd
seasonal rainfall

Grow19.20 <- input.daily[1:194,]
Grow19.20$dayssinceJul1 <- 1:194

Grow20.21 <- input.daily[195:559,]
Grow20.21$dayssinceJul1 <- 1:365

Grow21.22 <- input.daily[560:907,]
Grow21.22$dayssinceJul1 <-1:348

@@

#below is precip data from liaweenee - it only goes up to Jan 2022 right now
rainfall_L <- read.csv("R:/SET/PlantSci/Shared Work Spaces/Ecology/Peoples files/Marion Nyberg/PhD/Rainfall/rainfall.csv")
rainfall_L <- rainfall_L[-4]
colnames(rainfall_L)[4] <- "Precip"
rainfall_L$Year <- as.factor(rainfall_L$Year)
rainfall_L <- rainfall_L %>% filter(Year == 2019)
colnames(rainfall_L)[1] <- "jday"
colnames(rainfall_L)[3] <- "month"
rainfall_L$Site <- "L"
rainfall_L <- rainfall_L %>% filter(month >= 5)
rainfall_L$month <- as.numeric(rainfall_L$month) 
rainfall_L$month<- month.abb[rainfall_L$month]

@@

rainfall$season <- factor(rainfall$season , levels=c('EarlySummer','LateSummer', 'Autumn', 'Winter', 'Spring'))

input.daily$season <- "" # add season variable

# define seasons
winter <- (input.daily$month == "6" | input.daily$month == "7" | input.daily$month == "8")
spring <- (input.daily$month == "9" | input.daily$month == "10" | input.daily$month == "11")
earlysummer <- (input.daily$month == "12")
latesummer <- (input.daily$month == "1" | input.daily$month == "2")
autumn <- (input.daily$month == "3" | input.daily$month == "4" | input.daily$month == "5")

input.daily$season[winter] <- "Winter"
input.daily$season[spring] <- "Spring"
input.daily$season[earlysummer] <- "EarlySummer"
input.daily$season[latesummer] <- "LateSummer"
input.daily$season[autumn] <- "Autumn"

input.daily$season <- factor(input.daily$season , levels=c('EarlySummer','LateSummer', 'Autumn', 'Winter', 'Spring'))
```

```{r create rainfall summary}
rainfall.summary <- rainfall %>% ddply(c("Year", "Site", "month"), summarise,
                                       monthly.sum = sum(Precip, na.rm = TRUE))

Seasonal.rain <- rainfall%>% filter(Year == 2020 | Year == 2021 | Year == 2022) %>% ddply(c("Year", "season"), summarise,
                    seasonal.sum = sum(Precip, na.rm = TRUE))

rain.aov <- aov(seasonal.sum~Year*season, data =Seasonal.rain)
summary(rain.aov)

TukeyHSD(rain.aov)
```

@@

```{r cumulative fluxes, fig.width = 10, fig.height = 6}
cum.co2fluxes <- input.daily %>%
  select(NEE_LT.gC, GPP_LT.gC, ER_LT.gC, jday, Year, month) %>% 
  group_by(Year) %>% 
  mutate(NEEcum = cumsum(NEE_LT.gC),
         GPPcum = cumsum(GPP_LT.gC),
         ERcum = cumsum(ER_LT.gC))

(NEEcum.p <-cum.co2fluxes %>% 
  ggplot(., aes(jday, NEEcum, color = Year)) +
  geom_line(size = 1)+
  ylab(bquote('Cumulative NEE (g C' ~m^-2~')'))+
  xlab("Julian day of year")+
  theme_classic()+ 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.position = "none"))

(NEE.time <- input.daily %>% 
  ggplot(., aes(jday, NEE_LT.gC, color = Year)) +
  geom_line(size = 1)+
  ylab(bquote('Cumulative NEE (g C' ~m^-2~ day^-1*')'))+
  xlab("Julian day of year")+
  theme_classic()+ 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.position = "none"))

@@

(GPPcum.p <-cum.co2fluxes %>% 
  ggplot(., aes(jday, GPPcum, color = Year)) +
  geom_line(size = 1)+
  ylab(bquote('Cumulative GPP (g C' ~m^-2~')'))+
  xlab("Julian day of year")+
  theme_classic()+ 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.position = "none"))

(ERcum.p <-cum.co2fluxes %>% 
  ggplot(., aes(jday, ERcum, color = Year)) +
  geom_line(size =1)+
  ylab(bquote('Cumulative ER (g C' ~m^-2~')'))+
  xlab("Julian day of year")+
  theme_classic()+ 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.position = "none"))

@@

>>>>

Grow19.20 <- input.daily[1:194,]
Grow19.20$dayssinceJul1 <- 173:366
Grow19.20$GSYear <- "19/20"

Grow20.21 <- input.daily[195:559,]
Grow20.21$dayssinceJul1 <- 1:365
Grow20.21$GSYear <- "20/21"

Grow21.22 <- input.daily[560:907,]
Grow21.22$dayssinceJul1 <-1:348
Grow21.22$GSYear <- "21/22"

@@

#below is precip data from liaweenee - it only goes up to Jan 2022 right now
rainfall_L <- read.csv("R:/SET/PlantSci/Shared Work Spaces/Ecology/Peoples files/Marion Nyberg/PhD/Rainfall/rainfall.csv")
rainfall_L <- rainfall_L[-4]
colnames(rainfall_L)[4] <- "Precip"
rainfall_L$Year <- as.factor(rainfall_L$Year)
rainfall_L <- rainfall_L %>% filter(Year == 2019)
colnames(rainfall_L)[1] <- "jday"
colnames(rainfall_L)[3] <- "month"
rainfall_L$Site <- "L"
rainfall_L <- rainfall_L %>% filter(month >= 6)
rainfall_L$month <- as.numeric(rainfall_L$month) 
rainfall_L$month<- month.abb[rainfall_L$month]
rainfall_L$GSYear <- "19/20"

input.rearranged$season <- "" # add season variable

# define seasons
winter <- (rainfall_L$month == "Jun" | rainfall_L$month == "Jul" | rainfall_L$month == "Aug")
spring <- (rainfall_L$month == "Sep" | rainfall_L$month == "Oct" | rainfall_L$month == "Nov")
earlysummer <- (rainfall_L$month == "Dec")
latesummer <- (rainfall_L$month == "Jan" | rainfall_L$month == "Feb")
autumn <- (rainfall_L$month == "Mar" | rainfall_L$month == "Apr" | rainfall_L$month == "May")

rainfall_L$season[winter] <- "Winter"
rainfall_L$season[spring] <- "Spring"
rainfall_L$season[earlysummer] <- "EarlySummer"
rainfall_L$season[latesummer] <- "LateSummer"
rainfall_L$season[autumn] <- "Autumn"

rainfall_L$season <- factor(rainfall_L$season , levels=c('EarlySummer','LateSummer', 'Autumn', 'Winter', 'Spring'))

@@

rainfall$season <- factor(rainfall$season , levels=c('EarlySummer','LateSummer', 'Autumn', 'Winter', 'Spring'))
```

@@

```{r cumulative fluxes, fig.width = 10, fig.height = 6}
cum.co2fluxes <- input.rearranged %>% #input.daily %>%
  select(NEE_LT.gC, GPP_LT.gC, ER_LT.gC, dayssinceJul1, GSYear) %>% 
  group_by(GSYear) %>% 
  mutate(NEEcum = cumsum(NEE_LT.gC),
         GPPcum = cumsum(GPP_LT.gC),
         ERcum = cumsum(ER_LT.gC))

(NEEcum.p <-cum.co2fluxes %>% 
  ggplot(., aes(dayssinceJul1, NEEcum, color = GSYear)) +
  geom_line(size = 1)+
  ylab(bquote('Cumulative NEE (g C' ~m^-2~')'))+
  xlab("Days since July 1st")+
  theme_classic()+ 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13)))

(NEE.time <- input.daily %>% 
  ggplot(., aes(jday, NEE_LT.gC, color = Year)) +
  geom_line(size = 1)+
  ylab(bquote('Cumulative NEE (g C' ~m^-2~ day^-1*')'))+
  xlab("Days since July 1st")+
  theme_classic()+ 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.position = "none"))

@@

(GPPcum.p <-cum.co2fluxes %>% 
  ggplot(., aes(dayssinceJul1, GPPcum, color = GSYear)) +
  geom_line(size = 1)+
  ylab(bquote('Cumulative GPP (g C' ~m^-2~')'))+
  xlab("Days since July 1st")+
  theme_classic()+ 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13)))

(ERcum.p <-cum.co2fluxes %>% 
  ggplot(., aes(dayssinceJul1, ERcum, color = GSYear)) +
  geom_line(size =1)+
  ylab(bquote('Cumulative ER (g C' ~m^-2~')'))+
  xlab("Days since July 1st")+
  theme_classic()+ 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13)))

@@

```{r}
input.rearranged$season <- "" # add season variable

# define seasons
winter <- (input.rearranged$month == "Jun" | input.rearranged$month == "Jul" | input.rearranged$month == "Aug")
spring <- (input.rearranged$month == "Sep" | input.rearranged$month == "Oct" | input.rearranged$month == "Nov")
earlysummer <- (input.rearranged$month == "Dec")
latesummer <- (input.rearranged$month == "Jan" | input.rearranged$month == "Feb")
autumn <- (input.rearranged$month == "Mar" | input.rearranged$month == "Apr" | input.rearranged$month == "May")

input.rearranged$season[winter] <- "Winter"
input.rearranged$season[spring] <- "Spring"
input.rearranged$season[earlysummer] <- "EarlySummer"
input.rearranged$season[latesummer] <- "LateSummer"
input.rearranged$season[autumn] <- "Autumn"

input.rearranged$season <- factor(input.rearranged$season , levels=c('EarlySummer','LateSummer', 'Autumn', 'Winter', 'Spring'))
```

```{r}
rainfall.SP <- input.rearranged %>% select(Precip, month, GSYear, season)

rainfall <- rainfall_L %>% select(Precip, month, GSYear, season) %>% rbind(., rainfall.SP)

rainfall %>%  mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("GSYear", "season"), summarise,
  seasonal.total = sum(Precip)) %>%
  ggplot(., aes(x = GSYear, y = seasonal.total, group = GSYear, fill = season)) +
       #scale_x_discrete(limits = month.abb) +
  geom_col() + 
  ylab("Seasonal rainfall (mm)")+
  xlab("Month")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14))
```
